% arara: pdflatex: {files: [MathSACpr2014]}
% !arara: indent: {overwrite: yes}
\chapter{Program/Discipline Overview}

\epigraph{In two years, Portland Community College will be nationally known for
	progress from developmental math to college level courses and completion.}
{Chris Chairsell, Vice President of Academic and Student Affairs; Portland
	Community College Inservice, September 16, 2013}

\section[Educational objectives]{What are the educational goals or objectives of this
  program/discipline.
  How do these compare with national or professional program/discipline trends or guidelines?
  Have they changed since the last review, or are they expected to change in the next five years?
 }

As with any undergraduate or developmental education department, the primary goal of the faculty in the Math SAC can be summarized as follows: we hope to support students' life goals by imparting the skills and cognitive abilities necessary for continued success as they navigate their way through the education system and into the workforce.

As evidenced in the remainder of this document, we have an active faculty who are continually trying innovative strategies to achieve this goal.
Many of these strategies have been targeted directly at increasing student success and completion, such as: \begin{itemize} \item Accelerated Math Placement (AMP) and other placement enhancement tools (see \cpageref{other:sec:amp}); \item study-skills focused classroom activities (see \cpageref{cur:sub:studyskills}); \item experimentation with interactive homework/learning systems as well as development of said systems targeted to our students (see \cpageref{other:sec:webwork} and \cpageref{sec3:subset:alekspilot}); \item establishment and dissemination of best practices for online accessibility (see \cpageref{needs:sec:access}).
\end{itemize}

While each of these are worthy strategies, it has become increasingly apparent that something of greater scope needs to take place if we hope to see dramatic changes in the success and completion rates for students taking mathematics courses---especially those who initially place into developmental math courses.
The call for change nationwide in community colleges from access, to access and completion reinforces the Math SAC's awareness of the vital role we play in creating an environment of student success and completion.
We were pleased to hear Dr.
Chairsell call for the creation of a college-wide culture for math success, and we are very encouraged by the way in which this challenge has been embraced by departments such as student services.
We in the Math SAC also embrace Dr.
Chairsell's call for the creation of a college-wide culture for math success.
We are dedicated to making the necessary changes to our mathematics curriculum in order to maximize success and completion rates for our students.
By necessity, the most dramatic changes will need to take place in our developmental mathematics courses.
As we restructure, we are focused on integrating evidence-based best practices in order to achieve the highest rates of success and completion for our students.

\subsection{Developmental Education (DE)}
Historically viewed as `remediation', developmental education has often been marginalized by higher education entities.
However, practices in developmental education have been given wide scientific attention, and best practices are supported with extensive research.

The two largest organizations involved in developmental education research and professional development, the National Council of Developmental Education (NCDE) and the National Association of Developmental Education (NADE), define developmental education as \emph{a comprehensive process that focuses on the intellectual, social, and emotional growth and development of all students}.

The number of reasons students place into pre-college courses are too numerous to list.
However, a large number of students entering at the DE mathematics level have the added burden of an intense anxiety that hinders their ability to be successful in a mathematics course.
In combination with the academic, social, economic, and psychological issues facing students in DE math courses, we must approach any changes with the whole student in mind.

Over the last several years, there has been a growing sense that the traditional algebra content, as currently taught in our DE math courses, was not meeting the needs of many of our students.
The fact that in Fall 2013, over twenty-five SAC members joined the DE Math subcommittee formed specifically to take a deeper look at  the developmental math sequence is a strong indicator of the interest and concerns we hold.

\subsection{Science, Technology, Engineering, and Mathematics (STEM)}
Another emerging trend over the past five years has been a nationwide spotlight on STEM education and the dire need to increase the number of college students who ultimately obtain undergraduate degrees in STEM fields.
In fact, President Obama has formally designated increasing the number of undergraduate STEM majors by 1 million over the next decade as a Cross-Agency Priority (CAP) goal.
\footnote{\url{http://www.whitehouse.gov/blog/2012/12/18/one-decade-one-million-more-stem-graduates}}

The CAP goal proposes to focus efforts in five promising areas of opportunity: \begin{itemize} \item identifying and implementing evidence-based practices to improve STEM teaching and to attract students to STEM courses (see \cpageref{webworkposter,reflect:page:stem}); \item providing more opportunities for students to engage in meaningful  STEM activities through research experiences, especially in their first two years of college (see \cpageref{over:sec:DEfuture}); \item addressing the mathematics preparation gap that students face when they arrive at college, using evidence-based practices that generate improved results; \item providing educational opportunities and supports for women and historically underrepresented minorities; and \item identifying and supporting innovation in higher education.
\end{itemize}

The Math SAC realizes that for many students entering at the developmental education level, math courses serve as a barrier for those who might otherwise choose to pursue careers in STEM; this is well documented in studies such as PCAST: Engage to Excel \cite{engagetoexcel}.
As we work to recreate our developmental math curriculum, we are mindful of the need to reform our courses in such a way that they no longer serve as a barrier to the success of our students, and so that they also serve as a gateway to STEM careers for students who may have steered away from math in the past.
Most of the goals stated as CAP \emph{areas of opportunity} include elements that can be addressed in our courses, and we hope to create courses that are in alignment with attainment of those goals.

In doing this work, we have an eye not only on students who (eventually) pursue four-year STEM degrees, but we also have a focus on students enrolled in PCC's many CTE programs.
We are committed to creating courses that support success and completion for students enrolled in CTE programs.
Our courses must not only promote successful completion of the math course, but they must also impart skills that are specifically needed by the students in their CTE courses and ultimately in their chosen careers.

\subsection{The future of DE and undergraduate math at PCC}\label{over:sec:DEfuture}
While we are still in conversation, some themes have begun to emerge.
Preliminary discussions have transpired that might lead us to revamp our developmental education courses with an emphasis on the following: \begin{itemize} \item evidence-based best practices; \item streamlining the developmental education sequence; \item creating developmental education sequences which support STEM education and, ideally, promotes STEM education; \item integrating content into our developmental math courses that will create a math literate populace (intelligent consumers of data and problem solvers); \item tracking our progress through data-analysis and assessment that ensures that completion measures of  pass/fail rates do not mask a decrease in quality education.
\end{itemize}

While our current focus is on DE and STEM, we are also mindful of the need to reexamine our undergraduate level courses.
The content and teaching practices we adopt for our developmental  education courses need to be created with a clear understanding of the potential effects those changes will have on the students enrolled in our undergraduate level courses.
Additionally, we need to ensure that our commitment to using evidence-based best practices makes its way into the classroom for all of our courses, not just our developmental education courses.

We are excited by this opportunity to restructure our courses in ways that better support student success and completion.
We realize that this change cannot be developed or implemented in isolation and we look forward to discussing our ideas for DE restructure at the program review meeting.
We also look forward to continuing collaborative conversations with all stakeholders including, but not limited to, administration, CTE faculty, advising, counseling, testing, student services, union representatives, and---ultimately---the students themselves.

\subsubsection{Addendum}\label{over:subsub:nsfiuse}
In mid-December, 2013, members of the Math SAC were told that they were being given the opportunity to work with administration and members of a grant writing team to develop an NSF-IUSE grant proposal.
A team of four math faculty members was assembled for this work and over the next month they, along with the others working on the proposal, did a tremendous amount of research and met together at least weekly.
During this time we were working under the assumption that there were no restrictions of any kind on what we could propose in the grant.

A three-part model was developed that involved the creation of a math-specific CG course, a totally redesigned pathway structure through DE math, and a dramatic ramping-up of the way in which the computer platform WeBWorK would be used by our students.
The new pathways proposed would require that STEM-based guided learning activities be written for five new courses.
Our inspiration for the guided learning approach was based in part on work being done for the New Mathways Project,  which is a collaborative project between the Charles A.
Dana Center and the Texas Association of Community Colleges.
A diagram that illustrates the new pathway is shown in \cref{app:figure:pathways} (\vref{app:sec:pathways}).

On January 16, the day before our participation in the project was slated to end, the math team working on the grant was informed that in fact there were restrictions associated with the grant.
Because of the restrictions, the grant developers had cut the amount of person hours we had proposed by a factor of ten.
The newly proposed amount of faculty release was not nearly sufficient to achieve the goals we had proposed.
After much brainstorming about ways in which we might overcome that deficit or scale back the project, it was agreed by everyone involved---math faculty, administrators, and grant writers---that there was not a good fit between our proposal and that specific grant opportunity.

Not coincidentally, the Math SAC met on January \nth{17} in part to discuss the grant proposal.
At the end of that meeting the Math SAC, without a single no-vote, passed the following recommendations.

\recommendation[Grants Office, PCC Cabinet]{The Math SAC endorses the draft outline of the NSF-IUSE grant
	proposal, with the condition that the outline be fully supported with
	appropriate funding and services as determined by the Math SAC, and we encourage
	the Grants Office and PCC Cabinet to look into other funding sources.}

\recommendation[Math SAC, CG SAC, Completion Investment Council]{Regardless of the final design of the DE mathematics pathways,
	members of the Math SAC should continue to work with members of the CG SAC to
	create a comprehensive CG course designed specifically to address the needs of
	students registered in DE mathematics courses.}

\section[Changes since the last program review]{Please summarize changes that have been made since the last
  review.}
The mathematics department faculty is continually striving to improve our courses.
The recommendations from the 2003--2008 Program Review (PR) \cite{mathprogramreview2003} resulted in several changes as outlined in \vref{over:sec:changesresult}.
Some of the following changes that are mentioned in this section also appear at different sections of this document as referenced.
Most notably, our changes that are of a curricular nature are also addressed in  \vref{chap:otherissues}.

\subsection{Discontinue MTH 231 and MTH 232}
In Spring 2009 the Math SAC voted to discontinue offering MTH 231 and 232, our discrete mathematics courses.
The students taking these courses were mostly computer science students fulfilling requirements at PSU.
In order for the courses to transfer, the math department coordinated with the PCC and PSU computer science departments with respect to curriculum.
For various reasons it was mutually decided that the PCC computer science department should run the courses that are recognized statewide as CS 250 and 251.
\subsection{Formed the standing Math Learning Assessment Subcommittee
	(Math LAS)}
The committee was formed to address the college's assessment of the Core Outcomes.
\Cref{chap:outcomes} of this document (\cpageref{chap:outcomes})
outlines the results of this committee.
\subsection{Creation of MTH 84}
In Fall 2010 a pilot course was created to provide instruction in the use of the professional freeware publishing software \LaTeX.
While the emphasis of the course is creating professional mathematical documents, the skills learned can be used in a general context.
One online course was run each term and we received positive responses by students and faculty that took the course.
Students mention using the program in courses other than mathematics.
In May 2011 the Math SAC approved to make the one-credit course permanent (MTH 84) and we continue to run one online course every term---see \cpageref{other:sec:mth84} for more details.
\subsection{Creation of MTH 111H}
We approved the creation of a College Algebra honors course in Fall 2010.
A description can be found on \cpageref{cur:sub:111H}.
\subsection{Creation of faculty department co-chairs}
At the Cascade, Rock Creek and Sylvania campuses we offer between 100 to 150 class offerings per term, and thus we require a large part-time faculty pool to run these courses.
The formula used to measure the department chair load showed that each campus was either close to double if not more than double compared to the next highest faculty-chair load for any other discipline.
Starting in the Fall of 2010, the department chair positions at each of the mentioned campuses were split into co-chair positions.
Cascade, Rock Creek and Sylvania campuses each have two department co-chairs.
\subsection{Creation of SAC co-chairs option}
As the AY 2011/12 came to a close, the SAC voted to elect co-chairs for AY 2012/13, rather than a single SAC chair as had always been the custom.
This was repeated for AY 2013/14.
SAC by-laws will now be rewritten to allow the option for either a single SAC chair or SAC co-chairs.
The option for co-chairs may be most helpful in years surrounding Program Review, where SAC chair workload is higher than usual.
\subsection{Use of WeBWorK}
To further increase student accessibility and lower costs, Alex Jordan brought to our attention the freeware program, WeBWorK,  partially supported by  National Science Foundation grants.
The software is an online homework/testing system that can provide \emph{immediate} feedback to the student.
Spearheaded by Alex Jordan, faculty have been working on creating databases that fit our current curriculum.

We have been using WeBWorK since Spring 2009 and it is currently being used by several faculty in courses offered at PCC.
The advantages to the \emph{student} are that it is free and it is accessible to students with disabilities.
The advantages to \emph{faculty} are that we can adapt it to our own curriculum and can be used for other purposes besides coursework.
Ideas being proposed would allow for students to use it for preparation before taking placement exams.
We still are in the beginning phases as such a proposal would need to overcome technical difficulties.
Winter 2014 was the first term that we were able to run the program using PCC servers which allowed us to control the platform of this program.
Up to this point we had relied upon University of Oregon servers, which limited the capabilities of this program.
Further details are given on \cpageref{other:sec:webwork}.
\subsection{Social justice workgroup}
Four math faculty attended the conference ``Creating Balance in an Unjust World'' in San Francisco in Winter 2012 and were inspired to form an ongoing collaboration of faculty (including faculty from other disciplines) to create assignments and projects that have a social justice theme.
These faculty members share their assignments with others and encourage all math faculty to join them when they meet; see \cpageref{cur:sub:socialJustic} for more details.
\subsection{Credit hour change to MTH 243}
Students brought to our attention that our MTH 243 course was not transferring cleanly to some institutions.
To address this issue, MTH 243 changed from four to five credits effective Fall 2012.
An explanation for the change can be found in \vref{cur:sec:other}.
\subsection{Offering ALC math courses at Southeast}
To better serve students at Southeast, in Summer 2012, Southeast began offering self-paced basic math and introductory algebra courses (ALC Math) that were previously only offered at Sylvania.
\Vref{cur:sec:other}
contains further explanation of these courses.

\section[Changes resulting from the last program review]{Were any of the changes
  made as a result of the last review?
  If so, please describe the rationale and result.
 }\label{over:sec:changesresult}

In the 2003--2008 PR and the corresponding Administrative Response (AR), a large number of recommendations were given from the Math SAC and the Administration.
This section will look at changes that have been made due to those recommendations and some recommendations that are still being addressed.

\subsection{Incorporation of MTH 20 from DE}
One of the recommendations from the 2003--2008 PR (\cite{mathprogramreview2003}, page 30) was to transfer MTH 20 from the Developmental Education department to the Mathematics department.
That change has taken effect starting Fall 2013.
The change helped to align Sylvania with the rest of the campuses as to how this course was viewed.
Due to lack of resources (at other campuses) MTH 20 was, for all practical purposes, under the jurisdiction of the Math department.
Due to this change instructors teaching ALC math courses asked to also be incorporated into the Math department, housing all math courses under one legislative body.
The move was completed as of January 2013 (see \vref{app:sec:alc}).

\subsection{Alternative methods to accelerate completion}
The Administrative Response (AR) gave a list of recommendations (page 3) relating to alternative methods of moving students through the math sequence and accelerated math sequences.
In response to this recommendation, MTH 07/MTH 08 Accelerated Math Review (see \vref{other:sec:amp}) were created by the Math SAC.
Now that these classes are available we are hoping to offer more sections.
This will require more advertising when students are placed into a math class.
Additionally, since the ALC math classes have been moved into the Math SAC, the math faculty have become more aware of these courses.
The ALC classes were once only available at Sylvania, but now Southeast has incorporated the sequence and other campuses are looking into it.

\subsection{Assessment and course outcomes}
Page 2 of the AR asks the Math SAC to look at assessment more and take our Course Outcomes to the next level.
Please see \vref{chap:outcomes} (of this document) for details.
We have made major improvements on this front and have a standing assessment committee and action committee.
Some of our faculty members have roles in the college-wide assessment strategies and the Learning Assessment Council.

\subsection{Removal of MTH 91 and MTH 92}
The success rates for MTH 91 and 92 and additionally MTH 61, 62, and 63 were mentioned on page 2 of the AR.
After looking at the success rates of MTH 91 and 92, the Math SAC no longer offers these sections.
MTH 61, 62, and 63 are still being offered, but the Math SAC continues to work on Developmental Math and we currently have two committees looking at alternative Math Pathways.
Our hope is that a revised math curriculum could improve success and completion rates for students who usually take MTH 61, 62, 63.

\subsection{Study skills website}
Page 28 of the PR (\cite{mathprogramreview2003}) recommended that an orientation to `Studying at college' be part of the general orientation process.
Since the college has yet to make changes in this area, Jessica Bernards created study skills videos and activities that are currently being used by math faculty in Developmental Math Classes; further details are discussed on \cpageref{cur:sub:studyskills}.

\subsection{Enrollment issues with MTH 105}
A recommendation in the PR (\cite{mathprogramreview2003} page 31) wanted department chairs to look at MTH 105's low enrollment.
Since then MTH 111 B and C have been merged into a single MTH 111 class and the enrollment numbers in MTH 105 have increased.
Additionally two committees are currently looking at math pathways from the pre-college classes that might also increase 105 numbers.
This change also led to the adoption of a new MTH 111 textbook.
Normally changing a book wouldn't merit mention in a PR, but this book has a different philosophy and has therefore added additional changes to the college level math sequence.

\subsection{Distance learning standing committee}
A large number of recommendations from the last PR (\cite{mathprogramreview2003} pages 32--34) are related to Distance Learning.
We currently have a DL standing committee that looks into these matters.
See \vref{other:sec:distancelearning} (of this document) for a list of changes and concerns that the Distance Learning Standing Committee is currently working on.

\subsection{Academic interventions}
Finally, the last PR (\cite{mathprogramreview2003}, pages 30--31) made suggestions related to faculty contact with students outside the classroom and in the learning center.
The math faculty has continued to support the learning center over the last five years.
At Cascade Campus the math faculty have been working with retention specialists by creating academic interventions for students of concern.
This program shows promise and the specialists now have an office in the math department and three staff members.

