% arara: pdflatex: { files: [latexindent]}
\section{Introduction}
\subsection{Thanks}
	I first created \texttt{latexindent.pl} to help me format chapter files in a big project.
	After I blogged about it on the \TeX{} stack exchange \cite{cmhblog} I received some
	positive feedback and follow-up feature requests. A big thank you to Harish Kumar
	\cite{harish} who helped to develop and test the initial versions of the script.

	The \texttt{YAML}-based interface of \texttt{latexindent.pl} was inspired by the
	wonderful \texttt{arara} tool; any similarities are deliberate, and I hope that it is
	perceived as the compliment that it is. Thank you to Paulo Cereda and the team for
	releasing this awesome tool; I initially worried that I was going to have to make a GUI
	for \texttt{latexindent.pl}, but the release of \texttt{arara} has meant there is no
	need.

	There have been several contributors to the project so far (and hopefully more in the
	future!); thank you very much to the people detailed in \vref{sec:contributors} for their
	valued contributions, and thank you to those who report bugs and request features at
	\cite{latexindent-home}.

\subsection{License}
	\texttt{latexindent.pl} is free and open source, and it always will be; it
	is released under the GNU General Public License v3.0.

	Before you start using it on any important files, bear in mind that
	\texttt{latexindent.pl} has the option to overwrite your \texttt{.tex} files. It will
	always make at least one backup (you can choose how many it makes, see
	\cpageref{page:onlyonebackup}) but you should still be careful when using it. The script
	has been tested on many files, but there are some known limitations (see
	\cref{sec:knownlimitations}). You, the user, are responsible for ensuring that you
	maintain backups of your files before running \texttt{latexindent.pl} on them. I think it
	is important at this stage to restate an important part of the license here:
	\begin{quote}\itshape
		This program is distributed in the hope that it will be useful,
		but WITHOUT ANY WARRANTY; without even the implied warranty of
		MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. See the
		GNU General Public License for more details.
	\end{quote}
	There is certainly no malicious intent in releasing this script, and I do hope that it
	works as you expect it to; if it does not, please first of all make sure that you have
	the correct settings, and then feel free to let me know at \cite{latexindent-home} with a
	complete minimum working example as I would like to improve the code as much as possible.

	\begin{warning}
		Before you try the script on anything important (like your thesis), test it out on the
		sample files in the \texttt{test-case} directory \cite{latexindent-home}.
		\index{warning!be sure to test before use}
	\end{warning}

	\emph{If you have used any version 2.* of \texttt{latexindent.pl}, there
		are a few changes to the interface; see \vref{app:differences} and the comments
		throughout this document for details}.

\subsection{About this documentation}
	As you read through this documentation, you will see many listings; in this version of
	the documentation, there are a total of \totallstlistings. This may seem a lot, but I
	deem it necessary in presenting the various different options of \texttt{latexindent.pl}
	and the associated output that they are capable of producing.

	The different listings are presented using different styles:

	\begin{minipage}{.4\textwidth}
		\cmhlistingsfromfile{demonstrations/demo-tex.tex}{\texttt{demo-tex.tex}}{lst:demo-tex}
	\end{minipage}%
	\hfill
	\begin{minipage}{.4\textwidth}
		This type of listing is a \texttt{.tex} file.
	\end{minipage}%

	\begin{minipage}{.4\textwidth}
		\cmhlistingsfromfile[style=fileExtensionPreference]{../defaultSettings.yaml}[width=.8\linewidth,before=\centering,yaml-TCB]{\texttt{fileExtensionPreference}}{lst:fileExtensionPreference-demo}
	\end{minipage}%
	\hfill
	\begin{minipage}{.4\textwidth}
		This type of listing is a \texttt{.yaml} file; when you see line numbers given (as here)
		it means that the snippet is taken directly from \texttt{defaultSettings.yaml}, discussed in
		detail in \vref{sec:defuseloc}.
	\end{minipage}%

	\begin{minipage}{.55\textwidth}
		\cmhlistingsfromfile[style=modifylinebreaks]{../defaultSettings.yaml}[MLB-TCB,width=.85\linewidth,before=\centering]{\texttt{modifyLineBreaks}}{lst:modifylinebreaks-demo}
	\end{minipage}%
	\hfill
	\begin{minipage}{.4\textwidth}
		This type of listing is a \texttt{.yaml} file, but it will only
		be relevant when the \texttt{-m} switch is active; see \vref{sec:modifylinebreaks}
		for more details.
	\end{minipage}%

	\begin{minipage}{.55\textwidth}
		\cmhlistingsfromfile[style=replacements]{../defaultSettings.yaml}[replace-TCB,width=.85\linewidth,before=\centering]{\texttt{replacements}}{lst:replacements-demo}
	\end{minipage}%
	\hfill
	\begin{minipage}{.4\textwidth}
		This type of listing is a \texttt{.yaml} file, but it will only
		be relevant when the \texttt{-r} switch is active; see \vref{sec:replacements}
		for more details.
	\end{minipage}%

	% \begin{latexonly}
	You will occasionally see dates shown in the margin (for example, next to this
	paragraph!)%
	\announce{2017-06-25}{announce} which detail
	the date of the version in which the feature was implemented; the `N' stands for `new as
	of the date shown' and `U' stands for `updated as of the date shown'. If you see
	\stardemo, it means that the feature is either new (N) or updated (U) as of the release
	of the current version; if you see \stardemo\, attached to a listing, then it means that
	listing is new (N) or updated (U) as of the current version. If you have not read this
	document before (and even if you have!), then you can ignore every occurrence of the
	\stardemo; they are simply there to highlight new and updated features. The new and
	updated features in this documentation (\gitRel) are on the following pages:
	\listOfNewFeatures % \end{latexonly}

\subsection{Quick start}\label{sec:quickstart}
	If you'd like to get started with \texttt{latexindent.pl} then simply type
	\begin{commandshell}
latexindent.pl myfile.tex
\end{commandshell}
	from the command line. If you receive an error message such as that given in
	\cref{lst:poss-errors}, then you need to install the missing perl modules.
	\begin{cmhlistings}[style=tcblatex,language=Perl]{Possible error messages}{lst:poss-errors}
Can't locate File/HomeDir.pm in @INC (@INC contains: /Library/Perl/5.12/darwin-thread-multi-2level /Library/Perl/5.12 /Network/Library/Perl/5.12/darwin-thread-multi-2level /Network/Library/Perl/5.12 /Library/Perl/Updates/5.12.4/darwin-thread-multi-2level /Library/Perl/Updates/5.12.4 /System/Library/Perl/5.12/darwin-thread-multi-2level /System/Library/Perl/5.12 /System/Library/Perl/Extras/5.12/darwin-thread-multi-2level /System/Library/Perl/Extras/5.12 .) at helloworld.pl line 10.
BEGIN failed--compilation aborted at helloworld.pl line 10.
\end{cmhlistings}
	\texttt{latexindent.pl} ships with a script to help with this process; if you run the
	following script, you should be prompted to install the appropriate modules.
	\begin{commandshell}
perl latexindent-module-installer.pl
\end{commandshell}
	You might also like to see
	\href{https://stackoverflow.com/questions/19590042/error-cant-locate-file-homedir-pm-in-inc}{https://stackoverflow.com/questions/19590042/error-cant-locate-file-homedir-pm-in-inc},
	for example, as well as \vref{sec:requiredmodules}.

\subsection{A word about regular expressions}
	\index{regular expressions!a word about}
	As you read this documentation, you may encounter the term \emph{regular expressions}.
	I've tried to write this documentation in such a way so as to allow you to engage with
	them or not, as you prefer. This documentation is not designed to be a guide to regular
	expressions, and if you'd like to read about them, I recommend \cite{masteringregexp}.
